\documentclass[12pt]{article}
\usepackage[hcentering,bindingoffset=20mm]{geometry}
\usepackage{placeins}
%\usepackage[numbib]{tocbibind} use for ref in table of contents
\usepackage{rotating}
%\usepackage[natbibapa]{apacite}
%\usepackage[square,sort,comma,numbers]{natbib}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{tabularx}
\linespread{1.3}
\usepackage{gensymb}
\usepackage{longtable}
\usepackage{lscape}
\usepackage{url}
\addtolength{\textwidth}{2cm}
\addtolength{\hoffset}{-1cm}
\addtolength{\textheight}{2cm}
\addtolength{\voffset}{-1cm}
\setlength{\parindent}{0pt}
\title{Supporting information for 'Development of quantitative PCR assays for the detection and enumeration of toxic \emph{Gambierdiscus lapillus} and \emph{Gambierdiscus polynesiensis} (Gonyaulacales, Dinophyceae)'.}
\date{}
\begin{document}
\maketitle
\FloatBarrier
\begin{longtable}{ | p{2cm} | p{2cm} | p{3cm} | p{3.5cm} | p{3.5cm} | }
\caption{Screening of macroalgal samples for \emph{G. lapillus} and \emph{G. polynesiensis} and cell density estimates via qPCR. \emph{G. lapillus} cells numbers were modelled on the type strain HG7. \emph{G. polynesiensis} cell numbers were modelled on the strain CG14. N/D denotes not detected; N/A denotes not attempted due to loss of sample.}\\
\hline
\label{tbl:MacroalgaeTable}
\textbf{Sample ID}&\textbf{Spatial replicate}&\textbf{Macroalgal substrate}&\textbf{\textit{G. lapillus} cell number}&\textbf{\textit{G. polynesiensis} cell number}\\
\hline
1&A&\emph{Padina} sp.&N/D&N/D\\
\hline
1&B&\emph{Sargassum} sp.&10.55
&N/D\\
\hline
1&C&\emph{Padina} sp.&2.75
&N/D\\
\hline
2&A&\emph{Padina} sp.&N/D&N/D\\
\hline
2&B&\emph{Padina} sp. \& \emph{Chnoospora sp.}&4.33
&N/D\\
\hline
2&C&\emph{Padina} sp.&4.27
&N/D\\
\hline
3&A&\emph{Padina} sp.&0.62
&N/D\\
\hline
3&B&\emph{Padina} sp.&N/D&N/D\\
\hline
3&C&\emph{Padina} sp.&1.99
&N/D\\
\hline
4&A&\emph{Padina} sp. \& \emph{Chnoospora sp.}&6.13
&N/D\\
\hline
4&B&\emph{Chnoospora sp.}&0.62
&N/D\\
\hline
4&C&\emph{Padina} sp. \& \emph{Chnoospora sp.}&N/D&N/D\\
\hline
5&A&\emph{Padina} sp.&N/D&N/D\\
\hline
5&B&\emph{Padina} sp.&2.41
&N/D\\
\hline
5&C&\emph{Padina} sp.&3.81
&N/D\\
\hline
6&A&\emph{Chnoospora sp.}&1.12
&N/D\\
\hline
6&B&\emph{Padina} sp.&1.65
&N/D\\
\hline
6&C&\emph{Padina} sp.&N/D&N/D\\
\hline
7&A&\emph{Padina} sp.&9.35
&N/D\\
\hline
7&B&\emph{Padina} sp.&N/D&N/D\\
\hline
7&C&\emph{Padina} sp.&N/D&N/D\\
\hline
8&A&\emph{Chnoospora sp.}&N/D&N/D\\
\hline
8&B&\emph{Padina} sp.&1.67
&N/D\\
\hline
8&C&\emph{Padina} sp. \& \emph{Chnoospora sp.}&3.61
&N/D\\
\hline
9&A&\emph{Padina} sp. \& \emph{Chnoospora sp.}&N/D&N/D\\
\hline
9&B&\emph{Padina} sp.&1.69
&N/D\\
\hline
9&C&\emph{Padina} sp.&1.92
&N/D\\
\hline
10&A&\emph{Padina} sp. \& \emph{Chnoospora sp.}&12.6
&N/D\\
\hline
10&B&\emph{Padina} sp. \& \emph{Chnoospora sp.}&4.01
&N/D\\
\hline
10&C&\emph{Padina} sp.&N/D&N/D\\
\hline
11&A&\emph{Padina} sp. \& \emph{Sargassum} sp.&N/D&N/D\\
\hline
11&B&\emph{Padina} sp. \& \emph{Sargassum} sp.&0.26
&N/D\\
\hline
11&C&\emph{Padina} sp. \& \emph{Sargassum} sp.&1.29
&N/D\\
\hline
12&A&\emph{Chnoospora sp.}&N/D&N/D\\
\hline
12&B&\emph{Chnoospora sp.}&17.09
&N/D\\
\hline
12&C&\emph{Chnoospora sp.}&4.27
&N/D\\
\hline
13&A&\emph{Chnoospora sp.}&N/D&N/D\\
\hline
13&B&\emph{Padina} sp. \& \emph{Chnoospora sp.}&49.51
&N/D\\
\hline
13&C&\emph{Padina} sp. \& \emph{Chnoospora sp.}&18.58
&N/D\\
\hline
14&A&\emph{Padina} sp. \& \emph{Chnoospora sp.}&0.91
&N/D\\
\hline
14&B&\emph{Padina} sp. \& \emph{Chnoospora sp.}&N/D&N/D\\
\hline
14&C&\emph{Chnoospora sp.}&5.95
&N/D\\
\hline
15&A&\emph{Padina} sp. \& \emph{Chnoospora sp.}&2.01
&N/D\\
\hline
15&B&\emph{Chnoospora sp.}&4.89
&N/D\\
\hline
15&C&\emph{Chnoospora sp.}&N/D&N/D\\
\hline
16&A&\emph{Chnoospora sp.}&6.70&N/D\\
\hline
16&B&\emph{Chnoospora sp.}&8.83
&N/D\\
\hline
16&C&\emph{Chnoospora sp.}&3.08
&N/D\\
\hline
17&A&\emph{Chnoospora sp.}&2.58
&N/D\\
\hline
17&B&\emph{Chnoospora sp.}&9.39
&N/D\\
\hline
17&C&\emph{Padina} sp. \& \emph{Chnoospora sp.}&N/D&N/D\\
\hline
18&A&\emph{Chnoospora sp.}&0.02
&N/D\\
\hline
18&B&\emph{Chnoospora sp.}&N/D&N/D\\
\hline
18&C&\emph{Chnoospora sp.}&9.24
&N/D\\
\hline
19&A&\emph{Chnoospora sp.}&5.27
&N/D\\
\hline
19&B&\emph{Padina} sp. \& \emph{Chnoospora sp.}&48.46
&N/D\\
\hline
19&C&\emph{Padina} sp. \& \emph{Chnoospora sp.}&2.71
&N/D\\
\hline
20&A&\emph{Padina} sp. \& \emph{Chnoospora sp.}&2.81
&N/D\\
\hline
20&B&\emph{Padina} sp. \& \emph{Chnoospora sp.}&10.26
&N/D\\
\hline
20&C&\emph{Chnoospora sp.}&N/D&N/D\\
\hline
21&A&\emph{Padina} sp. \& \emph{Chnoospora sp.}&5.50
&N/D\\
\hline
21&B&\emph{Chnoospora sp.}&1.23
&N/D\\
\hline
21&C&\emph{Padina} sp.&10.32
&N/D\\
\hline
22&A&\emph{Chnoospora sp.}&N/D&N/D\\
\hline
22&B&\emph{Chnoospora sp.}&37.68
&N/D\\
\hline
22&C&\emph{Chnoospora sp.}&5.57
&N/D\\
\hline
23&A&\emph{Chnoospora sp.}&7.86
&N/D\\
\hline
23&B&\emph{Chnoospora sp.}&N/D&N/D\\
\hline
23&C&\emph{Chnoospora sp.}&3.14
&N/D\\
\hline
24&A&\emph{Padina} sp. \& \emph{Chnoospora sp.}&0.72
&N/D\\
\hline
24&B&\emph{Padina} sp. \& \emph{Chnoospora sp.}&0.58
&N/D\\
\hline
24&C&\emph{Padina} sp.&3.68
&N/D\\
\hline
25&A&\emph{Padina} sp.&N/D&N/D\\
\hline
25&B&\emph{Padina} sp.&N/D&N/D\\
\hline
25&C&\emph{Padina} sp.&N/D&N/D\\
\hline
26&A&\emph{Sargassum} sp.&N/D&N/D\\
\hline
26&B&\emph{Sargassum} sp.&0.19
&N/D\\
\hline
26&C&\emph{Sargassum} sp.&0.18
&N/D\\
\hline
27&A&\emph{Sargassum} sp.&N/D&N/D\\
\hline
27&B&\emph{Sargassum} sp.&2.11
&N/D\\
\hline
27&C&\emph{Sargassum} sp.&2.05
&N/D\\
\hline
28&A&\emph{Padina} sp.&7.17
&N/D\\
\hline
28&B&\emph{Padina} sp.&2.67
&N/D\\
\hline
28&C&\emph{Padina} sp.&8.64
&N/D\\
\hline
29&A&\emph{Chnoospora sp.}&1.24
&N/D\\
\hline
29&B&\emph{Chnoospora sp.}&5.90
&N/D\\
\hline
29&C&\emph{Chnoospora sp.}&N/D&N/D\\
\hline
30&A&\emph{Padina} sp. \& \emph{Sargassum} sp.&4.09
&N/D\\
\hline
30&B&\emph{Padina} sp. \& \emph{Sargassum} sp.&2.58
&N/D\\
\hline
30&C&\emph{Padina} sp. \& \emph{Sargassum} sp.&N/D&N/D\\
\hline
31&A&\emph{Sargassum} sp.&1.91
&N/D\\
\hline
31&B&\emph{Sargassum} sp.&2.90
&N/D\\
\hline
31&C&\emph{Sargassum} sp.&3.97
&N/D\\
\hline
32&A&\emph{Padina} sp. \& \emph{Sargassum} sp.&2.24
&N/D\\
\hline
32&B&\emph{Chnoospora sp.} \& \emph{Sargassum} sp.&1.36
&N/D\\
\hline
32&C&\emph{Padina} sp. \& \emph{Sargassum} sp.&2.00
&N/D\\
\hline
33&A&\emph{Padina} sp.&4.73
&N/D\\
\hline
33&B&\emph{Padina} sp. \& \emph{Chnoospora sp.}&0.07
&N/D\\
\hline
33&C&\emph{Padina} sp. \& \emph{Chnoospora sp.}&2.44
&N/D\\
\hline
\end{longtable}
\FloatBarrier
\FloatBarrier
\begin{table}
\caption{Screening of tow net samples for free floating \emph{G. lapillus} and \emph{G. polynesiensis} and cell density estimates via qPCR. N/D denotes not detected.}
\label{tbl:NetTable}
\begin{tabular}{ | p{4cm} | p{4cm} |p{4cm} | p{4cm} | }
\hline
\textbf{Sample ID}&\textbf{Location}&\textbf{\emph{G. lapillus}cell density}&\textbf{\emph{G. polynesiensis} cell density}\\
\hline
34&Shark Bay A&N/D&N/D\\ %?
\hline
35&Shark Bay B&N/D&N/D\\
\hline
36&Shark Bay D&1.24
&N/D\\
\hline
37&North beach D&0.70
&N/D\\
\hline
38&Jetty&N/D
&N/D\\
\hline
\end{tabular}
\end{table}
\FloatBarrier
\newpage
\end{document}